{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-02-11 thesis/landslide.tex
data race

\subsection{Kernel annotations}
\subsection{Bug identification}
\subsection{Data race analysis}
\label{sec:landslide-datarace}
\subsubsection{Limited Happens-Before}
\label{sec:landslide-lhb}
\subsubsection{Pure Happens-Before}
\label{sec:landslide-phb}
\subsection{Preemption point identification}
\label{sec:landslide-pps}
% TODO: talk about how w/in func and w/out func are processed
\subsection{State space estimation}
\label{sec:landslide-estimate}
\subsection{Dynamic Partial Order Reduction}
\label{sec:landslide-dpor}
\subsection{Iterative Context Bounding}
\label{sec:landslide-icb}
\subsection{Heuristic loop, deadlock, and synchronization detection}

>>>>

% not worth discussing?
%\subsection{Kernel annotations}
%\subsection{Bug identification}

\subsection{Preemption point identification}
\label{sec:landslide-pps}
% TODO: talk about how w/in func and w/out func are processed
\subsection{State space estimation}
\label{sec:landslide-estimate}
\subsection{Dynamic Partial Order Reduction}
\label{sec:landslide-dpor}

\subsection{Data race analysis}
\label{sec:landslide-datarace}

Whenever a memory conflict is identified for DPOR as described above,
the access pair's corresponding locksets and/or happens-before edges are checked to determine if it's also a data race.
Note the distinction: DPOR memory conflicts indicate that two thread transitions,
if reordered, could produce different behaviour, even if all accesses therein are adequately synchronized;
while a data race indicates furthermore that the two threads can be interleaved precisely at the moment of one or both accesses.
% nb. i think "only one of the accesses can be preempted at and have the other interleaved inside, but not the other way around,
% is a property of cli/sti involvement in kernel space only.

The core of the comparison is in {\tt check\_locksets()} in {\tt memory.c}.
It checks each DPOR memory conflict's locksets, for limited happens-before,
and happens-before edges, for pure happens-before
(\sect{\ref{sec:background-hb}}).

\subsubsection{Limited Happens-Before}
\label{sec:landslide-lhb}

Conditions 1, 2, and 4 defined in \sect{\ref{sec:background-hb}}, provided the Limited Happens-Before definition for \#4,
coincide exactly with the conditions to be considered a memory conflict under DPOR.
Hence all that remains to be checked is \#3, the set of locks held by each thread at the time of access.

Routines for recording lockset changes and computing set intersection are found in {\tt lockset.c}.
Apart from standard data structure manipulation,
one algorithmic point of note is that locks are distinguished by types in addition to address.
This allows (e.g.) mutexes stored as part of the implementation of semaphores to protect a different set of accesses than are protected by the semaphore they implement.

% TODO: talk abt free re malloc false positives

\subsubsection{Pure Happens-Before}
\label{sec:landslide-phb}

In Pure Happens-Before,
condition 4 is replaced with the traditional distributed systems notion of Happens-Before \cite{lamport-clocks}.
Landslide implements this via the vector clocks approach described by {\sc FastTrack} \cite{fasttrack}.
I refer the reader interested in the vector clock algorithm itself to the {\sc FastTrack} paper,
limiting discussion here to Landslide's corresponding implementation of each inference rule.

Despite being algorithmically worse in the size of the vector clocks,
I use the {\sc Djit+} rules for reads and writes rather than the {\tt FastTrack} ones;
because Landslide tests should be limited to few threads to manage the state space size,
$O(n)$ is usually in the single digits, so I optimize for code simplicity.

\begin{itemize}
	\item Reads and Writes (writes vs reads comparison already handled by DPOR) ({\tt memory.c})
	\begin{itemize}
		\item \textsc{Djit+ read/write same epoch} - {\tt vc\_eq()} case of {\tt add\_lockset\_to\_shm()}
		\item \textsc{Djit+ read/write} - {\tt vc\_happens\_before()} case of {\tt check\_locksets()}
	\end{itemize}
	\item Synchronization ({\tt schedule.c})
	\begin{itemize}
		\item \textsc{FT acquire}
			\begin{itemize}
				\item {\tt kern\_mutex\_\{,try\}locking\_done()} cases of {kern\_update\_state\_machine()}
				\item {\tt user\_mutex\_\{,try\}lock\_exiting()} cases of {user\_update\_state\_machine()}
				\item {\tt cli} case of {\tt kern\_update\_state\_machine()} (Pintos only)
				\item {\tt cli}/{\tt sti} lock handoff case in {\tt sched\_update()} (Pebbles only)
			\end{itemize}
		\item \textsc{FT release}
			\begin{itemize}
				\item {\tt kern\_mutex\_unlocking()} case of {kern\_update\_state\_machine()}
				\item {\tt user\_mutex\_unlock\_entering()} case of {user\_update\_state\_machine()}
				\item {\tt sti} case of {\tt kern\_update\_state\_machine()} (Pintos only)
				\item {\tt cli}/{\tt sti} lock handoff case in {\tt sched\_update()} (Pebbles only)
			\end{itemize}
		\item \textsc{FT fork} - {\tt agent\_fork()}
		\item \textsc{FT join} - {\tt sched\_unblock()} case of {\tt kern\_update\_state\_machine()} (Pebbles only)
	\end{itemize}
\end{itemize}

\subsection{Iterative Context Bounding}
\label{sec:landslide-icb}
\subsection{Heuristic loop, deadlock, and synchronization detection}
